---
title: "Linear Regression"
author: "Dinesh and Mahesh"
output: rmarkdown::html_vignette

#knit: (function(input, ...) {rmarkdown::render(input, output_file="D:\Linkoping University\Year1\Sem1\Period 1\Lab4_repo\vignettes\LR.html")})

#render pdf - pdf_document
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
This vignette summarizes functions of LR package to fit the linear regression model where computations are done using OLS (Ordinary Least Squares) and gives examples on how to use them.

```{r, include = FALSE, warning=FALSE, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE) # suppress title check

```


# RC implementation

## Included methods

These methods are included to provide analysis on data.

    * print()   prints out the coefficients and coefficient names
    * plot()    plots two plots using ggplot2 with Liu theme
    * resid()   returns the vector of residuals 
    * pred()    returns the predicted values yfit
    * coef()    returns the coefficients as a named vector

```{r echo=FALSE}
linreg <- setRefClass("linreg",
                      fields = list(formula="formula",
                                    data="data.frame",
                                    m_X = "matrix",
                                    m_Y="matrix",
                                    Xt="matrix",
                                    XtX="matrix",
                                    betaestimates = "matrix",
                                    yfit="matrix",
                                    residual="matrix",
                                    nparameters = "integer",
                                    dof="integer",
                                    residualt = "matrix",
                                    residualvariance="numeric",
                                    residualstd = "numeric",
                                    betavariance = "matrix",
                                    bb = "numeric",
                                    tvalues="matrix",
                                    pvalues= "matrix",
                                    standardizedresiduals ="matrix",
                                    sqrtstresiduals = "matrix",
                                    export_formula = "formula",
                                    export_data = "character"
                      ),
                      methods = list(
                        initialize = function(formula, data){
                          #Independent and dependent variables
                          m_X <<- model.matrix(formula, data)
                          m_Y <<- as.matrix(data[all.vars(formula)[1]])
                          # Transpose matrix X, and multiply + solve
                          Xt <<- t(m_X)
                          XtX <<- solve(Xt %*% m_X)
                          betaestimates <<- XtX %*% Xt %*% m_Y
                          # Estimate y
                          yfit <<- m_X %*% betaestimates
                          # Estimate residuals
                          residual <<- m_Y - yfit
                          # Determine degrees of freedom
                          nparameters <<- length(betaestimates)
                          dof <<- length(m_Y) - nparameters
                          # Variances
                          residualt <<- t(residual)
                          residualvariance <<- as.numeric((residualt %*% residual) / dof)
                          residualstd <<- sqrt(residualvariance)
                          betavariance <<- residualvariance * XtX
                          bb <<- diag(betavariance)
                          # t-values
                          tvalues <<- betaestimates/sqrt(bb)
                          # p-values
                          pvalues <<- 2 * pt(abs(tvalues), dof, lower.tail = FALSE)
                          # Standardized residuals for summary
                          standardizedresiduals <<- residual / sd(residual)
                          sqrtstresiduals <<- sqrt(abs(standardizedresiduals))
                          
                          # saving names
                          export_formula <<- formula
                          export_data <<- deparse(substitute(data))
                        },
                        
                         print = function() {
                          cat(paste("Call: \n"))
                          cat(paste("linreg(formula = ",format(export_formula), ", data = ", export_data, ")\n\n", sep = ""))
                          cat(paste("Coefficients:\n"))
                          coef <- structure(as.vector(betaestimates), names= row.names(betaestimates))
                          myPrint(coef)
                        },
                        
                        resid = function(){
                          return(as.vector(residual))
                        },
                        pred = function(){
                          return(yfit)
                        },
                        coef = function(){
                          vec <- as.vector(betaestimates)
                          names(vec) <- colnames(m_X)
                          return(vec)
                        },
                        
                        plot = function(){
                          library(ggplot2)
                          #library(ggThemeAssist)
                          theme <-  theme(
                            plot.background = element_rect(color = "black"),
                            panel.background = element_rect(fill = "white", color = NA),
                            panel.grid.major = element_line(color = "#1c1c19", size = 0.5),
                            panel.grid.major.x = element_blank(),
                            panel.grid.minor.x = element_blank(),
                            panel.grid.major.y = element_blank(),
                            panel.grid.minor.y = element_blank(),
                            axis.line = element_line(color = "#1c1c19", size = 0.5),
                            axis.text = element_text(color = "#1c1c19", size = 6),
                            axis.ticks = element_line(color = "#38ccd6", size = 0.5),
                            axis.title.x = element_text(color = "#38ccd6", size = 14,
                                                        face = "bold"),
                            axis.title.y = element_text(color = "#38ccd6", size = 14,
                                                        face = "bold"),
                            panel.grid.minor = element_line(color = "#1c1c19", size = 5),
                            plot.caption = element_text(size = 10, hjust =
                                                          0.5),
                            plot.margin = unit(c(1.2, 1.2, 1.2, 1.2), "cm"),
                            axis.text.x = element_text(size = 8),
                            axis.text.y = element_text(size = 8)
                          )
        
                          
                          plot1 <- ggplot(data.frame(yfit, residual), aes(y=residual, x=yfit)) + geom_point(shape=21, size=3, colour="black", fill="white")
                          plot1 <- plot1 + theme
                          
                          plot1 <- plot1 + stat_summary(fun.y=median, colour="red", geom="line", aes(group = 1))
                          plot1 <- plot1 + ggtitle("Residuals vs fitted") + xlab(paste("Fitted Values \n", "linreg(", format(export_formula), ")"))
                          plot2 <- ggplot(data.frame(yfit, sqrtstresiduals), aes(y=sqrtstresiduals, x=yfit)) + geom_point(alpha = 0.6, shape=21, size=3, colour="black", fill="white")
                          plot2 <- plot2 + theme
                          plot2 <- plot2 + stat_summary(fun.y=median, colour="red", geom="line", aes(group = 1))
                          plot2 <- plot2 + ggtitle("Scale-Location") + xlab(paste("Fitted Values \n", "linreg(", format(export_formula), ")"))
                          plot2 <- plot2 + scale_x_continuous(breaks = seq(0.0, 1.5, by= 0.5))
                          
                          plotlist <- list(plot1, plot2)
                          return(plotlist)
                        },
                        summary = function(){
                          "Prints the summary of linear regression model."
                          cat(paste("linreg(formula = ", format(export_formula), ", data = ", export_data, ") :\n\n ", sep = ""))
                          x <- setNames(as.data.frame(cbind(betaestimates,as.matrix(sqrt(bb)),tvalues, formatC(pvalues, format = "e", digits = 2), p_star_cal(pvalues))), c("Coefficients","Standard error","t values", "p values", ""))
                          myPrint(x)
                          cat(paste("\n\nResidual standard error: ", residualstd, " on ", dof, " degrees of freedom: ", sep = ""))
                        }
                      ))


myPrint = function(x, stripoff = FALSE) {
  print(x)
}

p_star_cal = function(p_value) {
  x <- ifelse(p_value > 0.1, " ",
              (ifelse(p_value > 0.05, " . ",
                      (ifelse(p_value > 0.01, "*",
                              (ifelse(p_value > 0.001, "**","***")))))))
  return(x)
}

```

## Data

To explore the basic functions of LR, we'll use ``` iris ```.This is perhaps the best known database to be found in the pattern recognition literature. The data set contains 3 classes of 50 instances each, where each class refers to a type of iris plant. One class is linearly separable from the other 2; the latter are not linearly separable from each other. you can see first 10 rows below:

```{r,echo=FALSE}
data(iris)
head(iris,10)
```           

## print()

Prints out the coefficients and coefficient names, similar as done by the lm class.

```{r,echo=FALSE}
mod_object <- linreg(formula=Petal.Length~Species, data = iris)
mod_object$print()
```

## plot()

This function plots two plots using ggplot2 with theme.

```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE, paged.print=TRUE, r,echo=FALSE}
mod_object$plot()
```

## resid()

This function returns the vector of residuals.

```{r,echo=FALSE}
mod_object$resid()
```

## pred()

It returns the  predicted values of yfit.

```{r,echo=FALSE}
mod_object$pred()
```

## coef()

This function returns the cofficients as a named vector

```{r,echo=FALSE}
mod_object$coef()
```

## summary()

This function returns a similar printout as printed for lm objects, but only need to present the coefficients with their standard error, t-value and p-value as well as the estimate of sigma and the degrees
of freedom in the model.

```{r,echo=FALSE}
mod_object$summary()
```

## Code

```{r ref.label=knitr::all_labels(), echo = T, eval = F}

```




